Paper 1: Instrument development

### 2024-09-03 Rmd/Paper1.Rmd
Added CFA

library(readODS)                            # read .ods-files. Kilde: https://github.com/ropensci/readODS/
library(readxl)                             # read .xlsx-files. Kilde: https://readxl.tidyverse.org/

library(pander)                             # print pretty tables
library(dplyr)                              # rename columns
library(tidyr)                              # pivot_longer

library(ggplot2)

# multicollinearity checks
library("corrplot")                         # multicollinearity 
library(car)                                # vif       
library("olsrr")                            # vif

library(psych)                              # for skew
library(mice)                               # for missing

@@

```{r remove_NMAR_TRELISQ}
# remove NMAR + NTNU (19)
data_TPACK_PST_noNMAR <-
  data_trelis_q_ren[-c(19,34,44,45,46,95,102),]

@@

@@

>>>>

### read files
library(readODS)                            # read .ods-files. Kilde: https://github.com/ropensci/readODS/
library(readxl)                             # read .xlsx-files. Kilde: https://readxl.tidyverse.org/

### presentation
library(pander)                             # print pretty tables
library(dplyr)                              # rename columns
library(tidyr)                              # pivot_longer
library(ggplot2)

### screening
library("corrplot")                         # multicollinearity library(car)                                # vif       
library("olsrr")                            # vif
library(mice)                               # for missing

### EFA
library(psych)                              # for skew and EFA

### CFA
library(lavaan)
#library(lavaanPlot)
library(semTools)                           # to find omega

@@

Choices:

* treat the data as interval or ordinal? (Watkins, 2018)
* missing data: use imputation (McNeish, 2017; Schafer and Graham, 2002)
* linearity-check: scatterplot, collinearity (Flora et al., 2012)
* normal or nonnormal distribution? Normality is important for test statistic. For nonnormal use Satorra-Bentler scaled chi-square and robust SE (Flora et al., 2012; McNeish, 2017). Nonnormal means skew>2 and kurtosis>9 (ref?). 
* aim for approximate fit, not exact fit (Flora et al., 2012; Weston and Gore, 2006)
* item selection: communalities (h2) - high (0.6 < h2 > 0.8) or wide (0.2 < h2 > 0.8)

"The swain and FAiR packages in R came perform the Swain correction and the FAiR package can also compute the Bartlett correction"

"Swain correction tends to provide the smallest correction but provides researchers with the most power to detect whether their model fits poorly"

yuan.bentler forutsetter adf-estimering. (McNeish, 2017) som krever at utvalgsstørrelser er minst p(p+1)/2, der p er antall items i modellen. Med p=12 blir det 78 respondenter, som er mer enn jeg har (p=106).

test="satorra.bentler"  # for robust test når ikke-normalfordelte data
test="yuan.bentler(.mplus)" (McNiesh, 2017: small samples? NB: adf)  
estimator="ULS"?  
se="robust.sem" # standard error  
information="observed" # ref. Schafer & Graham, 2002, p. 162

```{r remove_NMAR_TRELISQ}
# remove NMAR + NTNU (19)
CFA_TPACK_PST <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),]

### Spesifisering av modellene

Hensikt: bekrefte measurement model, altså indikatorenes kobling til de latente faktorene. (ser ikke på korrelasjoner her?)

```{r specify_models}

pseudo_1faktor <- '
PCK =~ CK123 + PCK2 + PCK5 + CK5 
TK =~ TK1b + TK2 + TK34 
TPACK =~ TPACK2 + TPACK4 + TCK4 + TPK4 + TCK1 + TPK1
'

tinker2 <- '
TPACK =~  TK1b + TK2 + TK34 + TPACK2 + TPACK4 + TCK4 + TPK4 + TCK1 + TPK1
PCK =~ CK123 + PCK2 + PCK5 + CK5
'

### Test av modellene

```{r help_functions_cfa}

### HELP FUNCTIONS
do_cfa <- function (cfaspec, cfadata) {
  # 7 faktor modell basert på EFA
# estimator og missing: Rhemtulla et al. (2012)
# ref.: https://statistics.ohlsen-web.de/multiple-imputation-with-mice/
# Predictive mean matching (McNeish, 2017)
TPACK_PST_0 <- cfa(cfaspec, data= cfadata, std.lv=TRUE,
                   #fixed.x=TRUE, 
                   auto.fix.single=TRUE, 
                   #auto.var=TRUE, auto.cov.lv.x=TRUE,
                  #estimator="MLR", 
                  missing="fiml", #robust.two.stage",
             # DWLS URL: https://lavaan.ugent.be/tutorial/est.html
             estimator="ML", #missing="pairwise", 
             #estimator="WLS",# missing="ml",
             #estimator="ULS", 
             #ordered=TRUE,
             #test="satorra.bentler", 
             test="yuan.bentler",
             information="observed",
             se="robust.huber.white") # robust.sem
             #se="boot", bootstrap=1000, # 10000 parallel=
             #parallel = "snow", ncpus = 4) # ref: https://github.com/yrosseel/lavaan/issues/23

return (TPACK_PST_0)
}

print_model <- function (cfamodel) {
# The argument standardized = TRUE augments the output with standardized parameter values. Two extra columns of standardized parameter values are printed. In the first column (labeled Std.lv), only the latent variables are standardized. In the second column (labeled Std.all), both latent and observed variables are standardized. The latter is often called the ‘completely standardized solution’. (ref: https://lavaan.ugent.be/tutorial/sem.html)
  print("summary")
  print(summary(cfamodel, standardized=TRUE, fit.measures=TRUE)) 
  print("estimates")
  # CI suggested by Cheung et al. (2023) for discriminant validity < 1 (one-tailed)
  print(parameterEstimates(cfamodel, rsquare = TRUE, level=0.90))
  print("modindices")
  print(modindices(cfamodel, sort=TRUE, maximum.number = 5))            # mi should be below 3.84

  # ref Rönkkö and Cho (2022)
#discriminantValidity(TPACK_PST_1, merge=TRUE)
# reliability, ref.: https://groups.google.com/g/lavaan/c/tPEPXQqQR48
#semTools::reliability(TPACK_PST_1)                         # deprecated
  print("reliability")
  rel <- compRelSEM(cfamodel, dropSingle = FALSE)
  print(rel)
  return (rel)   # for later comparison of reliability between models
}

#### tinker2

```{r estimate_9}
# egenfaktor
model9 <- do_cfa(tinker2, CFA_TPACK_PST)
rel9 <- print_model(model9)

#### pseudo_1faktor

```{r estimate_10}
# egenfaktor
model10 <- do_cfa(pseudo_1faktor, CFA_TPACK_PST)
rel10 <- print_model(model10)

### Oppsummering av fit

```{r CFA_results_comparison}
kolonnenavn <- c("AIC", "chi-square", "df", "p-value","RMSEA",
                 "RMSEA (90%) CI lower","RMSEA (90%) CI upper", 
                 "CFI","TLI", "SRMR") #, "chi2/df", 
                 #"chi-square scaled Swain", "p-value Swain", 
                 #"swain chi2/df")
anbefalte_verdier <- c("x", "x", "x", ">0.05", "<0.06(0.05-0.10)", "x", "x", ">0.95(0.9-0.95)", ">0.95", "<0.08(0.08-0.15)") #, "<2", "x", ">0.05", "<2") # Weston and Gore (2006); Gallagher and Brown (2013)

# skal jeg bruke scaled eller robust?
# bruk tre siffer for å kunne skille mellom CFI på 0.002 (discriminatory validity, ref. Shaffer et al. (2016))

#                  fitMeasures(TPACK_PST_0,"chisq.scaled")/
#                    fitMeasures(TPACK_PST_0,"df.scaled"),
#                80.35, 0.002, 80.35/fitMeasures(TPACK_PST_0,"df.scaled")),3)

fit9 <- round(fitMeasures(model9,
                  c("aic","chisq.scaled", "df.scaled",
                    "pvalue.scaled","rmsea.robust",
                    "rmsea.ci.lower.robust","rmsea.ci.upper.robust",
                    "cfi.robust","tli.robust","srmr")),3)
fit10 <- round(fitMeasures(model10,
                  c("aic","chisq.scaled", "df.scaled",
                    "pvalue.scaled","rmsea.robust",
                    "rmsea.ci.lower.robust","rmsea.ci.upper.robust",
                    "cfi.robust","tli.robust","srmr")),3)

radnavn <- c("anbefalt", "tinker2", "pseudo_1faktor")

fit_table <- rbind(anbefalte_verdier,fit9)
fit_table <- rbind(fit_table, fit10)

colnames(fit_table) <- kolonnenavn
row.names(fit_table) <- radnavn

pander(fit_table)
```

### Results

```{r covariance_matrix}

#CFA_TPACK_PST <- CFA_TPACK_PST - 3 # flytt nullpunkt
# covariance matrix / "TK1a","TK567"
pander(create_matrix(CFA_TPACK_PST[,c("CK123", "CK5", "PCK2" , "PCK5" , "TK1b", "TK2", "TK34", "TCK1","TCK4", "TPK1","TPK4","TPACK2","TPACK4")])) # ,"TPK4"
```

To calculate the students TPACK-score we would only calculate the average of all the items, i.e. an integrative TPACK-model, where all the elements contributes to the students TPACK.

```{r scores}

TK <- c("TK1b", "TK2", "TK34") # "TK1a","TK567"
PCK <- c("CK123", "PCK2" , "PCK5" , "CK5")
TPACK <- c("TPK1","TPK4","TPACK2","TCK1","TPACK4", "TCK4") # , "TPK4"

TPACK_PST_items <- list("TK" = TK,
                        "PCK" = PCK,
                        "TPACK" = TPACK) # everyone contributes til TPACK; c(TK,"TCK4",TPK, "CK123","PCK2","PCK5"))
df <- CFA_TPACK_PST[,unlist(TPACK_PST_items[1:3])]
describe(df)      # with missing (see n)? (na.rm=TRUE)
score_TPACK <- scoreItems(TPACK_PST_items,df)
#print(score_TPACK,option=FALSE)
describe(score_TPACK$scores)

@@

```{r omega_mle4}
omega0 <- omegaFromSem(model10)
omega0
#print(c(omega1$alpha, omega1$omega.tot)

# get correlation matrix of observed variables
cor_mat <- lavInspect(model10, what="cor.ov")
omega(cor_mat, nfactors=3)

omega(CFA_TPACK_PST[,c("CK123", "CK5", "PCK2" , "PCK5" , "TCK1","TCK4", "TPK1","TPK4","TPACK2","TPACK4", "TK1b", "TK2", "TK34")]) # ,"TPK4"
```

@@

I test for discriminatory and convergent validity following @ronkko_updated_2022

```{r}
# ref.: https://link.springer.com/article/10.1007/s10490-023-09871-y
source("../../scripts/measureQ.R")
```

```{r eval=T}
#measureQ(model9, CFA_TPACK_PST)

measureQ(model10, CFA_TPACK_PST)
```
